Optimal updating magnitude in adaptive flat-distribution-sampling simulations

### 2017-09-15 doc/invt.tex
lj.c estimate of Gamma_k for multiple-bin updating scheme

A commonly-used multiple-bin updating scheme in metadynamics
is the Gaussian updating scheme,
whose updating matrix can be characterized by
a rigid Gaussian function centered on the current bin
(with some boundary adjustment),
%to satisfy the constraints discussed in Sec. \ref{sec:updating-matrix},
see Fig. \ref{fig:mat}(a) for an example.
%
In terms of the continuous collective variable, $z = i \, \delta z$,
the eigenvalues of this scheme are given by\cite{bussi2006}
\begin{align}
  \lambda_k
  =
  \exp\left(
        -
        \frac{ \pi^2 \, \sigma_z^2 \, k^2 }
             { 2 \, \Delta z^2 }
      \right)
  ,
  %\notag
  \label{eq:lambda_Gaussian}
\end{align}
where $\sigma_z$ is the width of the Gaussian
and $\Delta z = n \, \delta z \, g/2$
is the effective span
with $g = 1$ for a periodic variable or $2$ otherwise
(cf. Appendix \ref{sec:Gaussian_math}).

@@

\begin{figure}[h]\centering
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=1.0\linewidth]{fig/mat.pdf}
  }
  \caption{
    \label{fig:mat}
    Updating matrices for two
    homogeneous updating schemes of
    a non-periodic variable ($n = 500$).
    Here $j$ is the current bin,
    $i$ is the bin where updating occurs,
    and $w_{ij}$ is the relative updating magnitude.
    %
    (a) The Gaussian updating scheme of width $\sigma = 20$ bins.
    (b) The homogeneous bandpass updating scheme of $K = 20$.
    %
    Both updating schemes
    can be characterized by a rigid window
    centered on the current bin, $j$,
    with some boundary adjustments.
  }
\end{figure}

@@

We will consider two special classes of bandpass updating schemes,
homogeneous and polynomial bandpass updating schemes.
%
%A summary of updating schemes is given in Fig. \ref{fig:scheme}.
%
First,
by applying this construction to the basis of cosine and sine functions,
we get a homogeneous bandpass updating scheme
(cf. Appendix \ref{sec:homo_bandpass}),
which resembles the Gaussian updating scheme,
but with long-range wiggles,
as shown in Fig. \ref{fig:mat}(b).
%
In Sec. \ref{sec:lj},
we will further compare
the homogeneous bandpass updating scheme
with the Gaussian updating scheme
in an example.
%
%While both updating scheme are able to deliver
%a smooth and differentiable potential of mean force,
%the bandpass updating scheme is superior in asymptotic convergence,
%because of the set of clear-cut spectrum of eigenvalues.
%Eq. \eqref{eq:lambda_bandpass},
%instead of a Gaussian one, Eq. \eqref{eq:lambda_Gaussian},
%
%\footnote{As an estimate of error,
%  we note that truncating a linear or quadratic function
%  at mode $K$ for a non-periodic variable
%  results in an error roughly proportional to
%  $|\max_i u_i - \min_i u_i|^2/K^3$.}
%
The other class of polynomial bandpass updating schemes will be discussed below.

@@

Upon convergence,
the parameters, $c_k^{(s)}$,
can be related to the PMF.
By Eqs. \eqref{eq:Vi_target} and \eqref{eq:rho_Gaussian}, we have
\begin{align*}
  u^{(s)}(z)
  %u(s, z)
  \approx
  \ln p(z)
  +
  \frac{ \bigl(z - z_c^{(s)} \bigr)^2 }
  { 2 \, \bigl( \sigma_z^{(s)} \bigr)^2 }
  +
  \mathrm{const.}
\end{align*}
%
Comparing this to Eq. \eqref{eq:u_age}, we get
\begin{align}
  \hat c_0^{(s)}
  &\approx \ln p\bigl(z_c^{(s)}\bigr) + \mathrm{const.}
  ,
  \label{eq:c0hat_limit} \\
  c_1^{(s)}
  &\approx \left. \frac{d \ln p(z) } { d z } \right|_{z = z_c^{(s)}}
  ,
  \label{eq:c1_limit} \\
  \frac{
    \sqrt 2 \, c_2^{(s)} - 1
  } { \bigl( \sigma_c^{(s)} \bigr)^2 }
  &\approx \left. \frac{d^2 \ln p(z) } { d z^2 } \right|_{z = z_c^{(s)}}
  .
  \label{eq:c2_limit}
\end{align}
%
Thus,
$\hat c_0^{(s)}$ as a function of $z_c^{(s)}$
approximates the negative PMF, $\ln p(z)$;\cite{maragliano2006, *abrams2008}
$c_1^{(s)}/\sigma_z^{(s)}$
approximates the mean force at $z = z_c^{(s)}$.
%
and $\sqrt{2} \, c_2^{(s)}$,
serving as the relative ``pressure''
to restrain the width of the distribution,
would exceed $1$ in a barrier region
where the PMF is locally concave,
i.e. $-d^2\ln p(z)/dz^2 < 0$.
%
%
We also have approximately\cite{park2007, *nguyen2013, *zhang2015st}
\begin{equation}
  \hat c_0^{(s+1)} - \hat c_0^{(s)}
  \approx
  \frac{1}{2}\left(
    \frac{ c_1^{(s+1)} } { \sigma_z^{(s+1)} }
    +
    \frac{ c_1^{(s)} } { \sigma_z^{(s)} }
  \right)
  \left(
  z_c^{(s+1)} - z_c^{(s)}
  \right)
  .
  \label{eq:c0_est}
\end{equation}
%
%and upon convergence, $\exp c_1^{(s)}$ should be proportional to
%the partition function,
%$\int e^{-\beta \, \Phi(\mathbf x)-u^{(s)}(z(\mathbf x),t)} \, d\mathbf x$,
%as in simulated tempering,
%where
%$\beta = 1/(k_B T)$ is the inverse temperature,
%$\mathbf x$ denotes the internal degrees of freedoms,
%and
%$\Phi(\mathbf x)$ is the unbiased potential energy.
%

@@

The results for the $L = 64$ case
after $10^5 L^2$ steps are shown
in Fig. \ref{fig:pt_hist}.
%
The Gaussian histograms were evenly spaced with similar widths as intended,
%and the overall energy histogram was roughly flat,
as shown in Fig. \ref{fig:pt_hist}(a).
%
We computed the density of states by Eq. \eqref{eq:WHAM},
and found the critical temperature to be $T_c = 0.70163$,
in agreement with a previous study\cite{wang2001pre}.
%
Since Gaussian umbrellas were sufficiently narrow,
we could readily verify
Eqs. \eqref{eq:c0hat_limit}-\eqref{eq:c2_limit}:
the logarithmic density of states, $\ln g(E)$,
was roughly captured
by $\hat c_0^{(s)}$ as a function of $E_c^{(s)}$,
as shown Fig. \ref{fig:pt_hist}(b);
%
$c_1^{(s)}/\sigma_E^{(s)}$ roughly representing
the local or statistical temperature, $\beta(E) = d\ln g(E)/dE$,
manifested the backbending behavior\cite{kim2006, *kim2007, kim2010}
typical in a first-order phase transition,
as shown in Fig. \ref{fig:pt_hist}(c);
%
$\sqrt 2 \, c_2^{(s)}$,
was indeed above $1.0$
in the unstable region,
where $\beta(E)$ is increasing or, equivalently,
$\ln g(E)$ is locally convex,
as shown in Fig. \ref{fig:pt_hist}(d).

\begin{figure}[h]\centering
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=1.0\linewidth]{fig/pt_hist.pdf}
  }
  \caption{
    \label{fig:pt_hist}
    Sampling using adaptive Gaussian ensembles
    on the $64 \times 64$ ten-state Potts model.
    %
    (a) Normalized histograms (solid lines)
    and the reweighted distribution
    at the critical temperature, $\beta_c$ (dashed line).
    %
    (b) $\hat c_0^{(s)} = c_0^{(s)} - c_2^{(s)}/\sqrt 2$
    against $E_c^{(s)}$ (circles)
    roughly agreed with the logarithmic density of states,
    $\ln g(E)$ (dashed line).
    In either case,
    we have deducted the linear contribution, $\beta_c \, E$,
    and applied a proper shift so as to normalize
    the distribution at $\beta_c$.
    %
    (c) $c_1^{(s)}/\sigma_E$ roughly corresponded to
    the statistical temperature, $\beta(E)$.
    %
    (d) $\sqrt 2 \, c_2^{(s)}$
    exceeded $1.0$ in the unstable region,
    where $\beta(E)$ is increasing.
    \note{When the new data come,
    run the \texttt{update.sh},
    copy the new value of the inverse critical temperature,
    and the value in \texttt{lns for normalization \dots [value] (lnz)}
    in the output.
    Use these values to update \texttt{bc} and \texttt{shiftlnz}
    in \texttt{pt\_hist.gp}, and run \texttt{gnuplot} again.  }%
  }
\end{figure}

@@

For a general multiple-bin updating scheme,
each mode of bias potential
corresponds to a CMA with a dynamic weight,
which is not necessarily the optimal equal weight.
%
Below we give a simple correction
to recover the equally-weighted average.
%
We will use capital letters to denote equally-weighted averages.
%e.g. for variable $y$,
%$Y(t) \equiv \frac{1}{t} \sum_{\tau=1}^t y(\tau)$.
%\begin{equation}
%  Y(t) \equiv \frac{1}{t} \sum_{\tau=1}^t x(\tau)
%  .
%  \notag
%  %\label{eq:timeav}
%\end{equation}
%
Then, by averaging both sides of Eq. \eqref{eq:uhatt}, we get
\begin{align}
  \hat U_i(t)
  &\approx
  U_i (t)
  + F_i(t) - 1
  \approx
  U_i(t)
  + \ln \frac{ H_i(t) } { \rho_i }
  .
  \label{eq:uhatav}
\end{align}
%
This correction serves as
a generalization of Eq. \eqref{eq:vcorr_equil}
for FDS simulations under
a variable bias potential.\footnote{We can also
  derive Eq. \eqref{eq:uhatav}
  by forming an independent estimator
  of the unbiased distribution\cite{marsili2006},
  which is, by Eq. \eqref{eq:Vi_target},
  $\hat p_i(t) \approx h_i(t) \, e^{ u_i(t) }$,
  assuming a proper shift of $u_i(t)$.
  Then after time averaging, we get
  $\rho_i e^{\hat U_i(t)} = \hat P_i(t) \approx H_i \, e^{ U_i(t) }$.}
%
However, it does not remove the systematic bias
caused by the adaptive updates,
and thus still requires a sufficiently small updating magnitude.

@@

Here we give a method of measuring
the integrals of the autocorrelation functions, $\Gamma_k$'s,
in an adaptive FDS simulation
under a constant updating magnitude.
%
This method has the advantage of not requiring
explicit computation of
the autocorrelation functions of the eigenmodes.

Because the difference between $u_i$ and $v_i$,
is a constant of time [cf. Eq. \eqref{eq:v_def}],
${\tilde u}_{k}$ and ${\tilde v}_{k}$
share the same variance.
%
%$\operatorname{var} {\tilde u}_{k} =
% \operatorname{var} {\tilde v}_{k} =
% \left\langle {\tilde v}_{k}^2 \right\rangle$,
%\begin{equation*}
%  \operatorname{var} {\tilde u}_{k}
%  =
%  .
%\end{equation*}
%
Then, from Eqs. \eqref{eq:xt_eql} and \eqref{eq:xt2_eql}, we have
%
\begin{equation}
  \operatorname{var} {\tilde u}_k
  =
  \operatorname{var} {\tilde v}_{k}
  \approx
  \frac{1}{2} \, \Gamma_k \, \lambda_k \, a_0
  %\left( 1 - e^{-2 \, \lambda_k \, a_0 \, T_0} \right)
  ,
\label{eq:varXt}
\end{equation}
%
for a sufficiently long period, $T_0$.
%where we have used the fact that the ensemble average of ${\tilde v}_{k}$
%is zero in the second step.
%
The variance, $\operatorname{var} \tilde u_k$, can be computed
by trajectory averaging in a long simulation.
%
Note that Eq. \eqref{eq:varXt} would fail for a near zero, $\lambda_k$,
and it is best used for the single-bin updating scheme.%\footnote{Alternatively,
%  we can split the trajectory into blocks,
%  compute $\Gamma_k$ from the histogram fluctuation of each block
%  using Eq. \eqref{eq:Fk2_Gammak},
%  and then average the results.}

@@

Consider an FDS simulation
under a nearly exact bias potential.
%
By Eq. \eqref{eq:sh_ave},
the fluctuation of the instantaneous histogram
is then dominated by the random part,
%
\begin{equation}
  f_{*i}(t) \approx \xi_{*i}(t)
  ,
  \label{eq:f_xi}
\end{equation}
and
%the same holds for the Fourier transform,
$\tilde f_k(t) \approx \tilde \xi_k(t).$
%
If we define the time average,
%
\begin{align*}
  \tilde F_k(T) \equiv \frac 1 T \sum_{t = 1}^T \tilde f_k(t)
  =\mathcal F\left[ \frac{ H_i(T) } { \rho_i } \right]_k
  ,
\end{align*}
%
for the histogram fluctuation of mode $k$,
then the total fluctuation is
%
\begin{align}
  F^2(T)
  &\equiv
  \sum_{k=1}^{n-1} \tilde F_k^2(T)
  =
  \sum_{i=1}^{n} \rho_i \, F_{*i}^2(T)
  =
  \sum_{i=1}^n
  \frac{ [ H_i(T) -\rho_i]^2 }{\rho_i}
  ,
  \notag
  %\label{eq:hfluc_def}
\end{align}
%
where we have used Eq. \eqref{eq:parseval} in the last step.
%
From Eq. \eqref{eq:f_xi}, we have
%$$\bigl\langle \tilde F_k^2(T) \bigr\rangle
%=
%\sum_{t=-\infty}^\infty
%\frac{
%\bigl\langle
%  \tilde f_k(t) \, \tilde f_k(0)
%\bigr\rangle } { T }
%\approx
%\sum_{t=-\infty}^\infty
%\frac{
%\bigl\langle
%  \tilde \xi_k(t) \, \tilde \xi_k(0)
%\bigr\rangle } { T }
%=
%\frac{ \Gamma_k } { T },$$
%
\begin{align}
  \bigl\langle \tilde F_k^2(T) \bigr\rangle
  &=
  \frac{1}{T}
  \sum_{t=-\infty}^\infty
  \bigl\langle
    \tilde f_k(t) \, \tilde f_k(0)
  \bigr\rangle
  \notag \\
  &\approx
  \frac{1}{T}
  \sum_{t=-\infty}^\infty
  \bigl\langle
    \tilde \xi_k(t) \, \tilde \xi_k(0)
  \bigr\rangle
  =
  \frac{ \Gamma_k } { T }
  ,
  \notag
  %\label{eq:Fk2_Gammak}
\end{align}
%
and the total histogram fluctuation is
\begin{equation}
  \bigl\langle F^2(T) \bigr\rangle
  =
  \sum_{k=1}^{n-1}
  \bigl\langle \tilde F_k^2(T) \bigr\rangle
  =
  \sum_{k=1}^{n-1}
  \frac{ \Gamma_k } { T }
  =
  \frac{ \Gamma } { T }
  ,
  \label{eq:F2sum}
\end{equation}
where we have used Eq. \eqref{eq:Gammak_sum}
in the last step.
%
We may relate the histogram fluctuation
to the minimal error under the single-bin updating scheme,
Eq. \eqref{eq:Emin_sbin}, as
  $\Err(T) \approx \bigl\langle F^2(T) \bigr\rangle$.
%we may rephrase the optimality of the inverse-time schedule
%as the inequality,
%\begin{equation}
%  \Err(T)
%  \approx
%  \bigl\langle F^2(T) \bigr\rangle
%  .
%  \notag
%  %\label{eq:fluc}
%\end{equation}
%and the histogram fluctuation
%serves as a good estimate of the final error.
%

>>>>

A commonly-used multiple-bin updating scheme in metadynamics
is the Gaussian updating scheme,
whose updating matrix can be characterized by
a rigid Gaussian function centered on the current bin
(with some boundary adjustment).
%to satisfy the constraints discussed in Sec. \ref{sec:updating-matrix},
%see Fig. \ref{fig:mat}(a) for an example.
%
In terms of the continuous collective variable, $z = i \, \delta z$,
the eigenvalues of this scheme are given by\cite{bussi2006}
\begin{align}
  \lambda_k
  =
  \exp\left(
        -
        \frac{ \pi^2 \, \sigma_z^2 \, k^2 }
             { 2 \, \Delta z^2 }
      \right)
  ,
  %\notag
  \label{eq:lambda_Gaussian}
\end{align}
where $\sigma_z$ is the width of the Gaussian
and $\Delta z = n \, \delta z \, g/2$
is the effective span
with $g = 1$ for a periodic variable or $2$ otherwise
(cf. Appendix \ref{sec:Gaussian_math}).

@@

%\begin{figure}[h]\centering
%  \makebox[\linewidth][c]{
%    \includegraphics[angle=0, width=1.0\linewidth]{fig/mat.pdf}
%  }
%  \caption{
%    \label{fig:mat}
%    Updating matrices for two
%    homogeneous updating schemes of
%    a non-periodic variable ($n = 500$).
%    Here $j$ is the current bin,
%    $i$ is the bin where updating occurs,
%    and $w_{ij}$ is the relative updating magnitude.
%    %
%    (a) The Gaussian updating scheme of width $\sigma = 20$ bins.
%    (b) The homogeneous bandpass updating scheme of $K = 20$.
%    %
%    Both updating schemes
%    can be characterized by a rigid window
%    centered on the current bin, $j$,
%    with some boundary adjustments.
%  }
%\end{figure}

@@

We will consider two special classes of bandpass updating schemes,
homogeneous and polynomial bandpass updating schemes.
%
%A summary of updating schemes is given in Fig. \ref{fig:scheme}.
%
First,
by applying this construction to the basis of cosine and sine functions,
we get a homogeneous bandpass updating scheme
(cf. Appendix \ref{sec:homo_bandpass}),
which resembles the Gaussian updating scheme,
but with long-range wiggles.
%as shown in Fig. \ref{fig:mat}(b).
%
In Sec. \ref{sec:lj},
we will further compare
the homogeneous bandpass updating scheme
with the Gaussian updating scheme
in an example.
%
%While both updating scheme are able to deliver
%a smooth and differentiable potential of mean force,
%the bandpass updating scheme is superior in asymptotic convergence,
%because of the set of clear-cut spectrum of eigenvalues.
%Eq. \eqref{eq:lambda_bandpass},
%instead of a Gaussian one, Eq. \eqref{eq:lambda_Gaussian},
%
%\footnote{As an estimate of error,
%  we note that truncating a linear or quadratic function
%  at mode $K$ for a non-periodic variable
%  results in an error roughly proportional to
%  $|\max_i u_i - \min_i u_i|^2/K^3$.}
%
The other class of polynomial bandpass updating schemes will be discussed below.

@@

Upon convergence,
the parameters, $c_k^{(s)}$,
can be related to the PMF.
By Eqs. \eqref{eq:Vi_target} and \eqref{eq:rho_Gaussian}, we have
\begin{align*}
  u^{(s)}(z)
  %u(s, z)
  \approx
  \ln p(z)
  +
  \frac{ \bigl(z - z_c^{(s)} \bigr)^2 }
  { 2 \, \bigl( \sigma_z^{(s)} \bigr)^2 }
  +
  \mathrm{const.}
\end{align*}
%
Comparing this to Eq. \eqref{eq:u_age}, we get
\begin{subequations}
\begin{align}
  \hat c_0^{(s)}
  &\approx \ln p\bigl(z_c^{(s)}\bigr) + \mathrm{const.}
  ,
  %\label{eq:c0hat_limit}
  \\
  c_1^{(s)}
  &\approx \left. \frac{d \ln p(z) } { d z } \right|_{z = z_c^{(s)}}
  ,
  %\label{eq:c1_limit}
  \\
  \frac{
    \sqrt 2 \, c_2^{(s)} - 1
  } { \bigl( \sigma_c^{(s)} \bigr)^2 }
  &\approx \left. \frac{d^2 \ln p(z) } { d z^2 } \right|_{z = z_c^{(s)}}
  .
  %\label{eq:c2_limit}
\end{align}
\label{eq:ck_limit}
\end{subequations}
%
Thus,
$\hat c_0^{(s)}$ as a function of $z_c^{(s)}$
approximates the negative PMF, $\ln p(z)$;\cite{maragliano2006, *abrams2008}
$c_1^{(s)}/\sigma_z^{(s)}$
approximates the mean force at $z = z_c^{(s)}$.
%
and $\sqrt{2} \, c_2^{(s)}$,
serving as the relative ``pressure''
to restrain the width of the distribution,
would exceed $1$ in a barrier region
where the PMF is locally concave,
i.e. $-d^2\ln p(z)/dz^2 < 0$.
%
%
We also have approximately\cite{park2007, *nguyen2013, *zhang2015st}
\begin{equation}
  \hat c_0^{(s+1)} - \hat c_0^{(s)}
  \approx
  \frac{1}{2}\left(
    \frac{ c_1^{(s+1)} } { \sigma_z^{(s+1)} }
    +
    \frac{ c_1^{(s)} } { \sigma_z^{(s)} }
  \right)
  \left(
  z_c^{(s+1)} - z_c^{(s)}
  \right)
  .
  \label{eq:c0_est}
\end{equation}
%
%and upon convergence, $\exp c_1^{(s)}$ should be proportional to
%the partition function,
%$\int e^{-\beta \, \Phi(\mathbf x)-u^{(s)}(z(\mathbf x),t)} \, d\mathbf x$,
%as in simulated tempering,
%where
%$\beta = 1/(k_B T)$ is the inverse temperature,
%$\mathbf x$ denotes the internal degrees of freedoms,
%and
%$\Phi(\mathbf x)$ is the unbiased potential energy.
%

@@

The results for the $L = 64$ case
after $10^5 L^2$ steps are shown
in Fig. \ref{fig:pt_hist}.
%
The Gaussian histograms were evenly spaced with similar widths as intended,
%and the overall energy histogram was roughly flat,
as shown in Fig. \ref{fig:pt_hist}(a).
%
We computed the density of states by Eq. \eqref{eq:WHAM},
and found the critical temperature to be $T_c = 0.70163$,
in agreement with a previous study\cite{wang2001pre}.
%
Since Gaussian umbrellas were sufficiently narrow,
we could readily verify
Eqs. \eqref{eq:ck_limit}:
the logarithmic density of states, $\ln g(E)$,
was roughly captured
by $\hat c_0^{(s)}$ as a function of $E_c^{(s)}$,
as shown Fig. \ref{fig:pt_hist}(b);
%
$c_1^{(s)}/\sigma_E^{(s)}$, roughly representing
the local or statistical temperature, $\beta(E) = d\ln g(E)/dE$,
manifested the backbending behavior\cite{kim2006, *kim2007, kim2010}
typical in a first-order phase transition,
as shown in Fig. \ref{fig:pt_hist}(c);
%
$\sqrt 2 \, c_2^{(s)}$,
was indeed above $1.0$
in the unstable region,
where $\beta(E)$ is increasing or, equivalently,
$\ln g(E)$ is locally convex,
as shown in Fig. \ref{fig:pt_hist}(d).

\begin{figure}[h]\centering
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=1.0\linewidth]{fig/pt_hist.pdf}
  }
  \caption{
    \label{fig:pt_hist}
    Sampling using adaptive Gaussian ensembles
    on the $64 \times 64$ ten-state Potts model.
    %
    (a) Normalized histograms.
    %
    (b) $\hat c_0^{(s)} = c_0^{(s)} - c_2^{(s)}/\sqrt 2$ 
    against $E_c^{(s)}$ (circles)
    roughly agreed with the logarithmic density of states,
    $\ln g(E)$ (dashed line).
    In either case,
    we have deducted the linear contribution, $\beta_c \, E$,
    and applied a proper shift so as to normalize
    the distribution at the critical temperature, $\beta_c$.
    %
    (c) $c_1^{(s)}/\sigma_E$ (circles), roughly corresponding to
    the statistical temperature, $\beta(E) = d\ln g(E)/dE$;
    $\beta_c$ (dashed line).
    %
    (d) $\sqrt 2 \, c_2^{(s)}$
    exceeded $1.0$ in the unstable region,
    where $\beta(E)$ is increasing.
    \note{When the new data come,
    run the \texttt{update.sh},
    copy the new value of the inverse critical temperature,
    and the value in \texttt{lns for normalization \dots [value] (lnz)}
    in the output.
    Use these values to update \texttt{bc} and \texttt{shiftlnz}
    in \texttt{pt\_hist.gp}, and run \texttt{gnuplot} again.  }%
  }
\end{figure}

@@

For a general multiple-bin updating scheme,
each mode of bias potential
corresponds to a CMA with a dynamic weight,
which is not necessarily the optimal equal weight.
%
Below we give a simple correction
to recover the equally-weighted average.
%
We will use capital letters to denote equally-weighted averages.
%e.g. for variable $y$,
%$Y(t) \equiv \frac{1}{t} \sum_{\tau=1}^t y(\tau)$.
%\begin{equation}
%  Y(t) \equiv \frac{1}{t} \sum_{\tau=1}^t x(\tau)
%  .
%  \notag
%  %\label{eq:timeav}
%\end{equation}
%
Then, by averaging both sides of Eq. \eqref{eq:uhatt}, we get
\begin{align}
  \hat U_i(t)
  &\approx
  U_i (t)
  + F_i(t) - 1
  \approx
  U_i(t)
  + \ln \frac{ H_i(t) } { \rho_i }
  .
  \label{eq:uhatav}
\end{align}
%
This correction serves as
a generalization of Eq. \eqref{eq:vcorr_equil}
for FDS simulations under
a variable bias potential.\footnote{We can also
  derive Eq. \eqref{eq:uhatav}
  by forming an independent estimator
  of the unbiased distribution\cite{marsili2006},
  which is, by Eq. \eqref{eq:Vi_target},
  $\hat p_i(t) \approx h_i(t) \, e^{ u_i(t) }$,
  assuming a proper shift of $u_i(t)$.
  Time averaging the equation yields
  $\rho_i \, e^{\hat U_i(t)} = \hat P_i(t) \approx H_i \, e^{ U_i(t) }$.}
%
However, it does not remove the systematic bias
caused by the adaptive updates,
and thus still requires a sufficiently small updating magnitude.

@@

Here we give two methods of estimating
the integrals of the autocorrelation functions, $\Gamma_k$'s,
in an adaptive FDS simulation
under a constant updating magnitude.
%
They have the advantage of not requiring
explicit computation of
the autocorrelation functions of the eigenmodes.

First, because the difference between $u_i$ and $v_i$,
is a constant of time [cf. Eq. \eqref{eq:v_def}],
${\tilde u}_{k}$ and ${\tilde v}_{k}$
share the same variance.
%
%$\operatorname{var} {\tilde u}_{k} =
% \operatorname{var} {\tilde v}_{k} =
% \left\langle {\tilde v}_{k}^2 \right\rangle$,
%\begin{equation*}
%  \operatorname{var} {\tilde u}_{k}
%  =
%  .
%\end{equation*}
%
Then, from Eqs. \eqref{eq:xt_eql} and \eqref{eq:xt2_eql}, we have
%
\begin{equation}
  \operatorname{var} {\tilde u}_k
  =
  \operatorname{var} {\tilde v}_{k}
  \approx
  \frac{1}{2} \, \Gamma_k \, \lambda_k \, a_0
  %\left( 1 - e^{-2 \, \lambda_k \, a_0 \, T_0} \right)
  ,
\label{eq:varXt}
\end{equation}
%
for a sufficiently long period, $T_0$.
%
We can compute the variance, $\operatorname{var} \tilde u_k$,
by trajectory averaging in a long simulation,
and use Eq. \eqref{eq:varXt} to estimate $\Gamma_k$.
%
However, this method would fail for a near zero, $\lambda_k$,
and it is best used for the single-bin updating scheme,
where $\lambda_k \equiv 1$.

To use the method in a simulation under a multiple-bin updating scheme,
we can construct a virtual bias potential, $u_i^\mathrm{vir}$,
that emulates the bias potential in a simulation
under the single-bin updating scheme.
%
This virtual bias potential can be constructed as (cf. Appendix \ref{sec:ave_us}),
\begin{equation}
  u_i^\mathrm{vir}(t+1)
  =
  u_i^\mathrm{vir}(t)
  +
  a_0 \, [\hat u_i(t) - u_i^\mathrm{vir}(t)]
  ,
  \notag
\end{equation}
%
where $\hat u_i(t)$ is given by Eq. \eqref{eq:uhatt}.
%
We can then apply Eq. \eqref{eq:varXt} to $u_i^\mathrm{vir}$
with the effective $\lambda_k \equiv 1$.

In this method, the updating magnitude, $a_0$,
needs to be sufficiently small, such that
$a_0 < 1/\left(\max_k \Gamma_k \right)$.

Alternatively, we can estimate $\Gamma_k$ using Eq. \eqref{eq:Fk2_Gammak}.
%
We can split the trajectory into blocks of $T_\mathrm{blk}$ steps,
compute the histogram fluctuation, $\tilde F_k(T_\mathrm{blk})$,
and estimate $\Gamma_k$ from the average of
$T_\mathrm{blk} \, \tilde F_k^2(T_\mathrm{blk})$
over the blocks.
%
For this method to work,
the trajectory block should be large enough
to cover at least a few autocorrelation times (which is roughly $\Gamma_k/2$),
but also sufficiently small
to avoid the artificial histogram flattening
caused by the frequent non-equilibrium updates.
Thus, we require $\max_k \Gamma_k < T_\mathrm{blk} < 1/a_0$
(We used $T_\mathrm{blk} = 0.1/a_0$.)

@@

Consider an FDS simulation
under a nearly exact bias potential.
%
By Eq. \eqref{eq:sh_ave},
the fluctuation of the instantaneous histogram
is then dominated by the random part,
%
\begin{equation}
  f_{*i}(t) \approx \xi_{*i}(t)
  ,
  \label{eq:f_xi}
\end{equation}
and
%the same holds for the Fourier transform,
$\tilde f_k(t) \approx \tilde \xi_k(t).$
%
If we define the time average,
%
\begin{align*}
  \tilde F_k(T) \equiv \frac 1 T \sum_{t = 1}^T \tilde f_k(t)
  =\mathcal F\left[ \frac{ H_i(T) } { \rho_i } \right]_k
  ,
\end{align*}
%
for the histogram fluctuation of mode $k$,
then the total fluctuation is
%
\begin{align}
  F^2(T)
  &\equiv
  \sum_{k=1}^{n-1} \tilde F_k^2(T)
  =
  \sum_{i=1}^{n} \rho_i \, F_{*i}^2(T)
  =
  \sum_{i=1}^n
  \frac{ [ H_i(T) -\rho_i]^2 }{\rho_i}
  ,
  \notag
  %\label{eq:hfluc_def}
\end{align}
%
where we have used Eq. \eqref{eq:parseval} in the last step.
%
From Eq. \eqref{eq:f_xi}, we have
%$$\bigl\langle \tilde F_k^2(T) \bigr\rangle
%=
%\sum_{t=-\infty}^\infty
%\frac{
%\bigl\langle
%  \tilde f_k(t) \, \tilde f_k(0)
%\bigr\rangle } { T }
%\approx
%\sum_{t=-\infty}^\infty
%\frac{
%\bigl\langle
%  \tilde \xi_k(t) \, \tilde \xi_k(0)
%\bigr\rangle } { T }
%=
%\frac{ \Gamma_k } { T },$$
%
\begin{align}
  \bigl\langle \tilde F_k^2(T) \bigr\rangle
  &=
  \frac{1}{T}
  \sum_{t=-\infty}^\infty
  \bigl\langle
    \tilde f_k(t) \, \tilde f_k(0)
  \bigr\rangle
  \notag \\
  &\approx
  \frac{1}{T}
  \sum_{t=-\infty}^\infty
  \bigl\langle
    \tilde \xi_k(t) \, \tilde \xi_k(0)
  \bigr\rangle
  =
  \frac{ \Gamma_k } { T }
  ,
  %\notag
  \label{eq:Fk2_Gammak}
\end{align}
%
and the total histogram fluctuation is
\begin{equation}
  \bigl\langle F^2(T) \bigr\rangle
  =
  \sum_{k=1}^{n-1}
  \bigl\langle \tilde F_k^2(T) \bigr\rangle
  =
  \sum_{k=1}^{n-1}
  \frac{ \Gamma_k } { T }
  =
  \frac{ \Gamma } { T }
  ,
  \label{eq:F2sum}
\end{equation}
where we have used Eq. \eqref{eq:Gammak_sum}
in the last step.
%
We may relate the histogram fluctuation
to the minimal error under the single-bin updating scheme,
Eq. \eqref{eq:Emin_sbin}, as
  $\Err(T) \approx \bigl\langle F^2(T) \bigr\rangle$.
%we may rephrase the optimality of the inverse-time schedule
%as the inequality,
%\begin{equation}
%  \Err(T)
%  \approx
%  \bigl\langle F^2(T) \bigr\rangle
%  .
%  \notag
%  %\label{eq:fluc}
%\end{equation}
%and the histogram fluctuation
%serves as a good estimate of the final error.
%
